### 2012-11-16 inst/examples/comparing-methods.Rmd
Attempted to fix the sequential method example.
Still incorrect.  See sequential-method.Rmd

```{r}
X <- seq(-5,5,len=50)
obs <- data.frame(x = c(-4, -3, -1,  0,  2, 3),
                  y = c(-2,  0,  1,  2, -1, -1))
#x <- -5:5
#obs <- data.frame(x = -5:5, y = sin(x) + rnorm(length(x),sd=.1))
l <- 1
sigma_n <- 0.8
```

@@

```{r}
  SE <- function(Xi,Xj, l) exp(-0.5 * (Xi - Xj) ^ 2 / l ^ 2)
  cov <- function(X, Y) outer(X, Y, SE, l) 
```

Cholesky method

@@

Direct method

@@

  Direct sequential method, avoids matrix inverse instability
``` {r}
ef <- numeric(length(X))
cf <- matrix(0, nrow=length(X), ncol=length(X))
for(i in 1:length(obs$x)){
  ef <- ef + cov(X, obs$x[i]) * obs$y[i] / as.numeric(cov(obs$x[i],obs$x[i]) + sigma_n^2)
  cf <- cf + cov(X, X) - cov(X, obs$x[i])  %*% cov(obs$x[i], X) / as.numeric(cov(obs$x[i],obs$x[i]) + sigma_n^2)
}
```

@@

kernlab method

@@

mlegp method

@@

``` {r}
require(reshape2)
df <- data.frame(x = X, Ef = Ef, Y = Y, y_p = y_p, ef = ef, y_m = y_m)
df <- melt(df, id = "x")
ggplot(df)+ geom_jitter(aes(x, value, color = variable)) + geom_point(data = obs, aes(x,y))
````

>>>>

```{r}
X <- seq(-5,5,len=50)
obs <- data.frame(x = c(-4, -3, -1,  0,  2, 3),
                  y = c(-2,  0,  1,  2, -1, -1))
#x <- -5:5
#obs <- data.frame(x = -5:5, y = sin(x) + rnorm(length(x),sd=.1))
l <- 1
sigma_n <- 0.
```

@@

```{r}
  SE <- function(Xi,Xj, l=1) exp(-0.5 * (Xi - Xj) ^ 2 / l ^ 2)
  cov <- function(X, Y) outer(X, Y, SE, l) 
```

### Cholesky method

@@

### Direct method

@@

### Direct sequential method, avoids matrix inverse instability

``` {r}
ef <- numeric(length(X))
cf <- matrix(0, nrow=length(X), ncol=length(X))

A <- as.numeric( cov(obs$x[1], obs$x[1]) )
mu <- obs$y[1]

for(i in 2:length(obs$x)){
  mu <- obs$y[i] + cov(obs$x[i], obs$x[1:(i-1)]) %*% (obs$y[1:(i-1)] - mu) / (A + sigma_n^2)
  A <- as.numeric( cov(obs$x[i], obs$x[i]) - cov(obs$x[i], obs$x[1:(i-1)]) %*% cov(obs$x[1:(i-1)], obs$x[i]) / (A + sigma_n^2) )
}

ef <- cov(X, obs$x) %*% (obs$y - mu) / A
cf <- cov(X,X) - cov(X, obs$x) %*% cov(obs$x, X) / A

```

@@

### kernlab method

@@

These are the same as `y_p`, explicitly showing the $E_y = K(x, x_i) \alpha$ multiplication

``` {r}
y_mult <-  kernelMult(kernelf(gp), X, xmatrix(gp), as.matrix(alpha(gp))) 
y_se <- sapply(X, function(x) cov(x, obs$x) %*% alpha(gp))
rbind(as.numeric(y_p), as.numeric(y_mult), as.numeric(y_se))[,1:10]
```

Which demonstrates the kernel trick `alpha = cov_xx_inv %*% obs$y`

### mlegp method

@@

``` {r}
require(reshape2)
df <- data.frame(x = X, direct = Ef, Cholesky = Y, kernlab = y_p, sequential = ef)
df <- melt(df, id = "x")
ggplot(df)+ geom_jitter(aes(x, value, color = variable)) + geom_point(data = obs, aes(x,y))
````
